NeuronUnit: Validation Tests for Neurophysiology Models

### 2013-11-04 NeurounitFrontiers/Paper.tex
Replaced URL citations with \url fields

\documentclass{frontiersSCNS}
%\documentclass{article}
%\usepackage{biblatex}
%\usepackage{amsmath}
%\usepackage{amssymb}
%\usepackage{times}
%\usepackage{hyperref}
%\usepackage{graphicx}
\usepackage{wrapfig}
\usepackage{listings}
%\usepackage[usenames,dvipsnames]{xcolor}
%\usepackage{setspace}
\usepackage{python_syntax}
%\usepackage{bibentry}
\usepackage{todonotes}
\usepackage{url}
\usepackage{lineno}
%\usepackage{natbib}
%\usepackage{caption}

@@

Test-driven methodologies have started to see success in neuroscience as well. 
Modeling competitions in neuroscience, for example, are typically organized around a collection of simple validation criteria, implemented as executable tests. 
These competitions continue to drive important advances and improve scientists' understanding of the relative merits of different models. 
For example, the quantitative single neuron modeling competition (QSNMC) \citep{jolivet_quantitative_2008} investigates the complexity-accuracy tradeoff among reduced models of excitable membranes; 
the ``Hopfield'' challenge \citep{hopfield_what_2000} tested techniques for generating neuronal network form given its function; 
the Neural Prediction Challenge sought the best stimulus reconstructions, given neuronal activity \citep{neural_prediction_url}; the Diadem challenge is advancing the art of neurite reconstruction \citep{diadem_url}; and 
examples from other subfields of biology abound \citep{dream_url}.
% \begin{figure}%{r}{0.6\textwidth}
% %\vspace{-38px}
% \includegraphics[scale=0.75]{sciunit_overview.pdf}
% \caption{Proposal overview. 
% \small{i) The \textit{SciUnit} framework can generate discipline-specific libraries for tests generation and model interfacing.  
% \textit{NeuronUnit} is proposed and described here.  Experimental data guides test development; model generation informs and is informed by these \textit{SciUnit} libraries.  
% Each \text{SciUnit} test suite repository on GitHub is indexed by the \textit{SciDash} web application.  
% Test suites for OSB and QSNMC (described in the text) are being built. 
% \textit{SciDash} automatically runs the test collections in each suite and publicly displays the results.}}
% %\vspace{-10px}
% \label{fig:sciunit_overview}
% \end{figure}
% \leavevmode
% \vspace{-0px}
\begin{figure}
\vspace{-38px}
\centering
\includegraphics[scale=0.6]{diagram1.pdf}
\vspace{-55px}
\caption{NeuronUnit overview. NeuronUnit is set of testing tools built upon the discipline-agnostic SciUnit framework. NeuronUnit can in principle test arbitrary neurophysiology models using arbitrary data but we provide here an example using models described in NeuroML as part of the Open Source Brain (OSB) project \citep{osb_url}, and single neuron electrophysiology data available as part of the NeuroElectro Project \citep{neuroelectro_url}.  Records of test results for various model/test combinations are accessible via SciDash, which indexes GitHub repositories of these records, models, and tests so they can be searched and filtered by the community.}  
\label{fig:sciunit_overview}
\end{figure}
\leavevmode
\vspace{-0px}

Each of these examples has leveraged \emph{ad hoc} infrastructure to support test generation. 
While the specific criteria used to evaluate models varies widely between disciplines in neuroscience, the underlying test-driven methodology has many common features that could be implemented once. 
Recognizing this, we developed a discipline-agnostic framework for developing scientific validation test suites called \textit{SciUnit} \citep{sciunit_url}. 
Here we describe \textit{NeuronUnit\-}, which builds upon \textit{SciUnit}, allowing neuroscientists to build \textit{SciUnit} tests that validate neurophysiology models against electrophysiological data.  We provide a concrete example pipeline, showing how models described using \text{NeuroML} and provided freely by the Open Source Brain project \citep{osb_url} can be test in fully automated fashion using published, curated data available through the \textit{NeuroElectro Project} \citep{neuroelectro_url}, leveraging facilities from the \textit{NeuroTools} library to extract relevant features of model output.  This is summarized in Figure \ref{fig:sciunit_overview}, which shows the relationships between the layers described here.
% In addition to this open testing toolkit we also provide a library of tests for models of spiking single neurons.

@@

%In many cases, models require no modifications to take the new tests because the same type of model output is being requested.
\subsection{Implementing a Validation Test in \textit{SciUnit}}
Fig. \ref{fig:rate_test} shows how a scientist would implement spike count tests such as the one described above using \textit{SciUnit}. 
A \textit{SciUnit} validation test is an {instance} (i.e. an object) of a Python class implementing the \verbx{sciunit.Test} interface (cf. line 1). 
Here, we show a class \verbx{SpikeCountTest} taking three \emph{parameters} in its constructor (constructors are named \verbx{__init__} in Python, lines 9-10). 
The meaning of each parameter along with a description of the goodness-of-fit metric used by the test is documented on lines 4-7. 
%For convenience, we also make use of functions provided by the popular NumPy\cite{numpy_url} and SciPy\cite{scipy_url} libraries, although these are not required by \textit{SciUnit}.  
To create a \emph{particular} spike count test, we instantiate this class with particular experimental observations. 
For example, given observations from hippocampal CA1 cells (not shown), we can instantiate a test as follows\todo{use iPython style In[0] notation?}:
\begin{python}
  CA1_sc_test = SpikeCountTest(CA1_inputs, CA1_means, CA1_stds)
\end{python}
We emphasize the crucial distinction between the \textit{class} \verbx{SpikeCountTest}, which defines a \emph{parameterized family} of validation tests, and the particular \textit{instance} \verbx{CA1_sc_test}, which is an individual validation test because the necessary parameters, derived from data, have been provided. 
As we will describe below, we expect communities to build repositories of such families capturing the criteria used in their subfields of neuroscience so that test generation for a particular system of interest will often require simply instantiating a previously-developed family with particular experimental parameters and data. 
For single-neuron test families like \verbx{SpikeCountTest}, we have developed such a library, called \textit{NeuronUnit} \citep{neurounit_url} (Sec. \ref{sec:neuronunit}).

Classes that implement the \verbx{sciunit.Test} interface must contain a \verbx{_judge} method that receives a candidate \emph{model} as input and produces a \textit{score} as output. 
To specify the interface between the test and the model (that is, to specify an appropriate scope), the test author provides a list of \emph{capabilities} in the \verbx{required_capabilities} attribute, seen on line 12 of Fig. \ref{fig:rate_test}. 
Capabilities are simply collections of methods that a test may need to invoke to receive relevant data, and are analogous to \emph{interfaces} in e.g. Java \citep{java_interfaces_url}. 
In Python, they are written as classes with unimplemented members. 
The capability required by the test in Fig. \ref{fig:rate_test} is shown in Fig. \ref{fig:capability}. 
In \textit{SciUnit}, classes defining capabilities are tagged as such by inheriting from \verbx{sciunit.Capability}. The test in Figure \ref{fig:rate_test} uses this capability on line 19 to produce a spike count prediction for each input current.

@@

\subsection{\textit{NeuronUnit}: A Suite of Tests for Neurophysiology}\label{sec:neuronunit}
By using machine-readable models from Open Source Brain (OSB), and machine-readable data from resources like The NeuroElectro Project \citep{neuroelectro_url}, we have produced a suite of data-driven validation tests in \textit{NeuronUnit}, as well as tools for capability implementation, and model interfaces for a representative set of canonical neurophysiology models, all publicly available.

\label{sec:neuronunit_acitivities}
Here we describe standard neuroinformatics tools we have adopted to develop \textit{NeuronUnit} \citep{neurounit_url}.

@@

NeuroConstruct \citep{neuroconstruct_url,gleeson_neuroconstruct:_2007} is a simulation manager that takes NeuroML models and hands off simulation to supported simulators. 
\textit{NeuronUnit} offers a \verbx{sciunit.Model} subclass called \verbx{NeuroConstructModel}, instantiated with the path to a NeuroML model.  
Because NeuroML can describe such a wide range of models, \verbx{NeuroConstructModel} makes few assumptions about them: that each one is \verbx{TimeIntegrable}, and \verbx{HasMembranePotential}.  
It is subclassed to test \textit{specific} NeuroML models (Fig. \ref{fig:ca1_model}).

The Open Source Brain project (OSB, \citep{osb_url}) curates many models described in NeuroML. 
OSB-curated projects are converted from their native format into NeuroML, and run on major neural simulators \citep{neuron_url,genesis_url,nest_url,moose_url}. 
Concordance between model output (beginning with the NeuroML description) and reference output (from native simulator source files) is reported for each model. 
Thus, OSB is an excellent source of models that, in addition to being open source, are sufficiently described to enable validation. 
The hippocampal CA1 pyramidal cell is commonly modeled, and we implement one such model hosted on OSB \citep{osb_ca1_url} by simply declaring a \verbx{CA1PyramidalCellModel} class, inheriting from \verbx{NeuroConstructModel}.  
This basic implementation simply ``wraps'' the components of the existing model, with simulator interaction taken care of by \verbx{NeuroConstructModel} methods; 
thus, only the code shown in Fig. \ref{fig:ca1_model} is required. 
All OSB models, and indeed any NeuroML model, can be tested similarly.

Spanning a range of scales and original development environments, all OSB models are formally described using NeuroML, as are all model components and sub-components, such as cells, ion channels, calcium stores, etc. 
These models are regularly executed on OSB servers to ensure that their output remains consistent as they are updated. 
Therefore, OSB can confirm that they \textit{do} work, while linked journal articles, on-site wiki, and code inspection can establish \textit{how} they work. 
However, there is no mechanism for establishing \textit{how well} they work, i.e. how well the models accord with data. 
\textit{NeuroUnit} fills this gap by helping OSB (and the larger biology community) assess models using data-driven tests. 
\textit{SciUnit} can be applied similarly to other neuroscience (or biology) sub-disciplines using \textit{NeuronUnit} analogues written by the corresponding communities.

@@

\subsubsection{Capabilities from NeuroTools}
NeuroTools \citep{neuralensemble_url} is a Python library supporting tasks associated with analysis of neural data (or model output), such as membrane potential time series, spike trains, etc. 
It is an open source and actively developed project, containing reliable algorithms on which to base neurophysiology tests.

@@

\subsubsection{Reference Data for Tests from NeuroElectro}
Answering \textit{how well} requires validation testing against data. 
The NeuroElectro project \citep{neuroelectro_url} is an effort to curate all published single cell neurophysiology data \citep{tripathy_neuroelectro:_2012}.  
Currently, up to 27 electrophysiological properties are reported for 93 cell types, spanning $>$ 2000 single pieces of published data extracted from article tables. 
We have made it easy to construct \textit{NeuronUnit} tests using the NeuroElectro API to get reference data. 
Tests can be based upon data from single journal articles, or from ensembles of articles with a common theme (e.g. about a particular neuron type). 
The former is illustrated in Figure \ref{fig:neuronunit_example}. 
Associated statistics of that data (e.g mean, standard error, and sample size) are attached and enable judgement of model output according to a chosen scoring mechanism. 
While NeuroElectro alone cannot judge all model aspects, it can serve to validate basic features of many neurophysiology models, such as resting membrane potential, action potential width, after-hyperpolarization amplitude, etc. 
As NeuroElectro is the only publicly curated source of such data, it represents a key component for \textit{NeuronUnit} test constuction.

\subsection{\textit{SciDash}: A community web application}\label{sec:scidash}
A collection of tests, models, and test records can summarize the current state of modeling in a research area.  
Community-oriented cyberinfrastructure to support the creation and summarization of such collections is essential.
We used a service called \textit{SciDash} \citep{scidash_url} for coordinated development of software repositories.

@@

\textit{SciDash} exists in three layers.
The \textbf{first layer} is completely under the control of individual developer communities, and consists of git repositories \citep{ram_git_2013} stored on GitHub \citep{github_url}. 
\textit{SciDash} identifies all similarly structured repositories (by tracking the lineage of a vanilla \textit{SciDash} repository above using the GitHub API).
This list of repositories is indexed by \textit{SciDash} to form the \textbf{second layer}, an overview of the state of all \textit{SciDash} repositories on GitHub.  
This index is searchable and filterable on the \textit{SciDash} website to identify repositories of interest to a particular research community.  
The \textbf{third layer} is a ``dashboard'' view of the state of a field, visible through a matrix of records for model/test combinations in a GitHub repository (Fig. \ref{fig:scidash_repo}). 
Test records are visualized as a ``record matrix'' composed of large numbers of model/test combinations (Fig. \ref{fig:scidash_matrix}).  
Each row in this matrix contains results for all tests taken by one model and serve as clear evidence of that model's scope and validity.  
Hyperlinks in the record matrix point to models, tests, and records available on GitHub. 
Each record links to test records (stored in the repository), displaying line-by-line the executed code and intermediate output statements, as well as the underlying models and tests.

@@

\subsection{A Complete Pipeline}
Although the tools described in Sec. \ref{sec:neuronunit_acitivities} do not exhaust the possible sources of models, capabilities, and test data, they provide an immediate point of entry into the neurophysiology community and a powerful demonstration of our proposal. 
In the \textit{NeuronUnit} repository \citep{neurounit_url} is a runnable script (\textit{examples.py}) demonstrating a complete testing pipeline. 
It (1) selects an OSB model; 
(2) simulates it using NeuroConstruct; 
(3) tests the widths of the resulting action potentials, extracted and computed using NeuroTools, against NeuroElectro data downloaded on-the-fly, using a \textit{NeuronUnit} test class called \verbx{SpikeWidthTestDynamic}; 
and (4) computes and prints a test score (Figure~\ref{fig:neuronunit_example}).

@@

\subsection{Participation from Modeling Communities}
Modelers may not want to expose model capabilities, a requirement of test-taking.  
We anticipate four solutions: \textbf{First}, interfacing a model to \textit{SciUnit} requires only implementing selected model capabilities. 
Often this means identifying native model procedures that satisfy a capability, and wrapping their functionality. 
This can require as little as one line of code. 
Importantly, the modeler is not required to expose or rewrite any model flow control. 
\textbf{Second}, we support multiple environments automatically by using NeuroML \citep{gleeson_neuroml:_2010}, and other simulator-independent model descriptions are possible for other domains. 
Automated generation of NeuroML from native model source code is in development (Gleeson, personal communication); for the popular NEURON simulator \citep{neuron_url}, this functionality is already mature and in use. 
This minimizes modeler effort for a large and growing number of models. 
\textbf{Third}, modelers have an incentive to demonstrate publicly their models' validity. 
Participation in public modeling competitions demonstrates this incentive. 
\textbf{Fourth}, modelers have an incentive to use \textit{SciUnit} during development (see TDD, above) to ensure that ongoing development preserves correspondence between model and data. 
A popular test suite can represent a ``gold standard'' by which progress during development is judged.

@@

\subsection{Diversity of Levels and Kinds of Models and Data}
The diversity of topics in biology is vast. 
\textbf{First}, we address this by providing an interface allowing modelers to express specific capabilities.  
This capability set determines the range of eligible tests.  Scale hierarchies are embedded in capability inheritance.  =
For example, \verbx{HasActionPotentials} inherits from \verbx{HasNeurons}, and \verbx{HodgkinHuxley} inherits from \verbx{VoltageGated}. 
Thus, incompatibility of a test-requiring-action-potentials for a model-lacking-neurons is known without explicit tagging. 
\textbf{Second}, NeuroML naturally addresses diversity of scales because it is organized hierarchically, in ``levels.''  
Models can be sub- or supersets of other models; similarly for SBML \citep{sbml_url,hucka_systems_2003}, a general systems biology markup language. 
\textbf{Third}, cross-level testing can use ``Representional Similarity Analysis'' (RSA) \citep{kriegeskorte_representational_2008}, requiring only that a model respond to defined inputs (e.g. stimuli). 
A ``similarity matrix'' for input responses defines a unique model signature, and can serve as intermediate test output. 
Goodness-of-fit between similarity matrices for model and experiment determines test scores; 
these matrices are independent of model scale because their size depends only on test inputs, not system detail.

\subsection{Arbitrary Scoring Criteria for Tests}
A test first assesses goodness-of-fit, and applies a normalization (e.g. pass/fail, 0.0-1.0) to generate a score. 
Arbitrary choices at both stages may benefit some models over others.  \textbf{First}, however, rank-ordering is constant across many goodness-of-fit metrics, meaning that choice of metric will rarely cause an otherwise passing model to fail and vice versa. 
For example, given a data mean and variance, ordering model output by Z-score or p-value will yield the same relative ranking of models. 
Indeed, rank ordering of models may prove more valuable than test scores themselves. 
\textbf{Second}, suite repositories are open (e.g. Fig. \ref{fig:scidash_repo}), so tests can be cloned and new statistical choices implemented. 
Statistics as applied to neuroscience have been notoriously ``fast and loose''; 
identification and correction of flawed methodology is becoming increasingly common \citep{button_power_2013,kriegeskorte_circular_2009,galbraith_study_2010,fish_fmri_url}, and is accelerated by an open framework. 
The validation criteria is made explicit in the specification of a test, so modelers need not guess which criteria are being used to validate or invalidate their model. 
Validation criteria are subject to debate (indeed, the QSNMC criteria changed between 2007 and 2008 due to such debates), and neuroscientists who wish to promote different criteria need only derive alternative tests. 
A community consensus will slowly emerge in the form a commonly-accepted test suite (see Sec. \ref{sec:scidash}). 
This community can judge which test version is most appropriate, i.e. what a model \textit{should} do -- this process documented via standard moderation techiniques used on GitHub -- and the \textit{SciUnit} framework determines whether the model \textit{does} it.

>>>>

\documentclass{frontiersSCNS}
%\documentclass{article}
%\usepackage{biblatex}
%\usepackage{amsmath}
%\usepackage{amssymb}
%\usepackage{times}
\usepackage{hyperref}
%\usepackage{graphicx}
\usepackage{wrapfig}
\usepackage{listings}
%\usepackage[usenames,dvipsnames]{xcolor}
%\usepackage{setspace}
\usepackage{python_syntax}
%\usepackage{bibentry}
\usepackage{todonotes}
\usepackage{url}
\usepackage{lineno}
%\usepackage{natbib}
%\usepackage{caption}

@@

Test-driven methodologies have started to see success in neuroscience as well. 
Modeling competitions in neuroscience, for example, are typically organized around a collection of simple validation criteria, implemented as executable tests. 
These competitions continue to drive important advances and improve scientists' understanding of the relative merits of different models. 
For example, the quantitative single neuron modeling competition (QSNMC) \citep{jolivet_quantitative_2008} investigates the complexity-accuracy tradeoff among reduced models of excitable membranes; 
the ``Hopfield'' challenge \citep{hopfield_what_2000} tested techniques for generating neuronal network form given its function; 
the Neural Prediction Challenge sought the best stimulus reconstructions, given neuronal activity (\url{http://neuralprediction.berkeley.edu}); the Diadem challenge is advancing the art of neurite reconstruction (\url{http://www.diademchallenge.org}); and 
examples from other subfields of biology abound (\url{http://www.the-dream-project.org}).
% \begin{figure}%{r}{0.6\textwidth}
% %\vspace{-38px}
% \includegraphics[scale=0.75]{sciunit_overview.pdf}
% \caption{Proposal overview. 
% \small{i) The \textit{SciUnit} framework can generate discipline-specific libraries for tests generation and model interfacing.  
% \textit{NeuronUnit} is proposed and described here.  Experimental data guides test development; model generation informs and is informed by these \textit{SciUnit} libraries.  
% Each \text{SciUnit} test suite repository on GitHub is indexed by the \textit{SciDash} web application.  
% Test suites for OSB and QSNMC (described in the text) are being built. 
% \textit{SciDash} automatically runs the test collections in each suite and publicly displays the results.}}
% %\vspace{-10px}
% \label{fig:sciunit_overview}
% \end{figure}
% \leavevmode
% \vspace{-0px}
\begin{figure}
\vspace{-38px}
\centering
\includegraphics[scale=0.6]{diagram1.pdf}
\vspace{-55px}
\caption{NeuronUnit overview. NeuronUnit is set of testing tools built upon the discipline-agnostic SciUnit framework. 
NeuronUnit can in principle test arbitrary neurophysiology models using arbitrary data but we provide here an example using models described in NeuroML as part of the Open Source Brain (OSB) project (\url{http://www.opensourcebrain.org}), and single neuron electrophysiology data available as part of the NeuroElectro Project (\url{http://neuroelectro.org}).  
Records of test results for various model/test combinations are accessible via SciDash, which indexes GitHub repositories of these records, models, and tests so they can be searched and filtered by the community.}  
\label{fig:sciunit_overview}
\end{figure}
\leavevmode
\vspace{-0px}

Each of these examples has leveraged \emph{ad hoc} infrastructure to support test generation. 
While the specific criteria used to evaluate models varies widely between disciplines in neuroscience, the underlying test-driven methodology has many common features that could be implemented once. 
Recognizing this, we developed a discipline-agnostic framework for developing scientific validation test suites called \textit{SciUnit} (\url{http://www.sciunit.org}). 
Here we describe \textit{NeuronUnit\-}, which builds upon \textit{SciUnit}, allowing neuroscientists to build \textit{SciUnit} tests that validate neurophysiology models against electrophysiological data.  We provide a concrete example pipeline, showing how models described using \text{NeuroML} and provided freely by the Open Source Brain project (\url{http://www.opensourcebrain.org}) can be test in fully automated fashion using published, curated data available through the \textit{NeuroElectro Project} (\url{http://neuroelectro.org}), leveraging facilities from the \textit{NeuroTools} library to extract relevant features of model output.  This is summarized in Figure \ref{fig:sciunit_overview}, which shows the relationships between the layers described here.
% In addition to this open testing toolkit we also provide a library of tests for models of spiking single neurons.

@@

%In many cases, models require no modifications to take the new tests because the same type of model output is being requested.
\subsection{Implementing a Validation Test in \textit{SciUnit}}
Fig. \ref{fig:rate_test} shows how a scientist would implement spike count tests such as the one described above using \textit{SciUnit}. 
A \textit{SciUnit} validation test is an {instance} (i.e. an object) of a Python class implementing the \verbx{sciunit.Test} interface (cf. line 1). 
Here, we show a class \verbx{SpikeCountTest} taking three \emph{parameters} in its constructor (constructors are named \verbx{__init__} in Python, lines 9-10). 
The meaning of each parameter along with a description of the goodness-of-fit metric used by the test is documented on lines 4-7. 
%For convenience, we also make use of functions provided by the popular NumPy\cite{numpy_url} and SciPy\cite{scipy_url} libraries, although these are not required by \textit{SciUnit}.  
To create a \emph{particular} spike count test, we instantiate this class with particular experimental observations. 
For example, given observations from hippocampal CA1 cells (not shown), we can instantiate a test as follows\todo{use iPython style In[0] notation?}:
\begin{python}
  CA1_sc_test = SpikeCountTest(CA1_inputs, CA1_means, CA1_stds)
\end{python}
We emphasize the crucial distinction between the \textit{class} \verbx{SpikeCountTest}, which defines a \emph{parameterized family} of validation tests, and the particular \textit{instance} \verbx{CA1_sc_test}, which is an individual validation test because the necessary parameters, derived from data, have been provided. 
As we will describe below, we expect communities to build repositories of such families capturing the criteria used in their subfields of neuroscience so that test generation for a particular system of interest will often require simply instantiating a previously-developed family with particular experimental parameters and data. 
For single-neuron test families like \verbx{SpikeCountTest}, we have developed such a library, called \textit{NeuronUnit} (\url{http://github.com/scidash/neuronunit}) (Sec. \ref{sec:neuronunit}).

Classes that implement the \verbx{sciunit.Test} interface must contain a \verbx{_judge} method that receives a candidate \emph{model} as input and produces a \textit{score} as output. 
To specify the interface between the test and the model (that is, to specify an appropriate scope), the test author provides a list of \emph{capabilities} in the \verbx{required_capabilities} attribute, seen on line 12 of Fig. \ref{fig:rate_test}. 
Capabilities are simply collections of methods that a test may need to invoke to receive relevant data, and are analogous to \emph{interfaces} in e.g. Java (\url{http://docs.oracle.com/javase/tutorial/java/concepts/interface.html}). 
In Python, they are written as classes with unimplemented members. 
The capability required by the test in Fig. \ref{fig:rate_test} is shown in Fig. \ref{fig:capability}. 
In \textit{SciUnit}, classes defining capabilities are tagged as such by inheriting from \verbx{sciunit.Capability}. The test in Figure \ref{fig:rate_test} uses this capability on line 19 to produce a spike count prediction for each input current.

@@

\subsection{\textit{NeuronUnit}: A Suite of Tests for Neurophysiology}\label{sec:neuronunit}
By using machine-readable models from Open Source Brain (OSB), and machine-readable data from resources like The NeuroElectro Project (\url{http://www.neuroelectro.org}), we have produced a suite of data-driven validation tests in \textit{NeuronUnit}, as well as tools for capability implementation, and model interfaces for a representative set of canonical neurophysiology models, all publicly available.

\label{sec:neuronunit_acitivities}
Here we describe standard neuroinformatics tools we have adopted to develop \textit{NeuronUnit} (\url{http://github.com/scidash/neuronunit}) \todo{Must change name of repository on github to neuronunit (after I am sure grant has been reviewed)}.

@@

NeuroConstruct (\citep{gleeson_neuroconstruct:_2007}, \url{http://www.neuroconstruct.org}) is a simulation manager that takes NeuroML models and hands off simulation to supported simulators. 
\textit{NeuronUnit} offers a \verbx{sciunit.Model} subclass called \verbx{NeuroConstructModel}, instantiated with the path to a NeuroML model.  
Because NeuroML can describe such a wide range of models, \verbx{NeuroConstructModel} makes few assumptions about them: that each one is \verbx{TimeIntegrable}, and \verbx{HasMembranePotential}.  
It is subclassed to test \textit{specific} NeuroML models (Fig. \ref{fig:ca1_model}).

The Open Source Brain project (OSB, \url{http://www.opensourcebrain.org}) curates many models described in NeuroML. 
OSB-curated projects are converted from their native format into NeuroML, and run on major neural simulators (\url{http://www.neuron.yale.edu/neuron}, \url{http://genesis-sim.org}, \url{http://www.nest-initiative.org}, \url{http://moose.ncbs.res.in}). 
Concordance between model output (beginning with the NeuroML description) and reference output (from native simulator source files) is reported for each model. 
Thus, OSB is an excellent source of models that, in addition to being open source, are sufficiently described to enable validation. 
The hippocampal CA1 pyramidal cell is commonly modeled, and we implement one such model hosted on OSB (\url{http://opensourcebrain.org/projects/ca1pyramidalcell}) by simply declaring a \verbx{CA1PyramidalCellModel} class, inheriting from \verbx{NeuroConstructModel}.  
This basic implementation simply ``wraps'' the components of the existing model, with simulator interaction taken care of by \verbx{NeuroConstructModel} methods; 
thus, only the code shown in Fig. \ref{fig:ca1_model} is required. 
All OSB models, and indeed any NeuroML model, can be tested similarly.

Spanning a range of scales and original development environments, all OSB models are formally described using NeuroML, as are all model components and sub-components, such as cells, ion channels, calcium stores, etc. 
These models are regularly executed on OSB servers to ensure that their output remains consistent as they are updated. 
Therefore, OSB can confirm that they \textit{do} work, while linked journal articles, on-site wiki, and code inspection can establish \textit{how} they work. 
However, there is no mechanism for establishing \textit{how well} they work, i.e. how well the models accord with data. 
\textit{NeuronUnit} fills this gap by helping OSB (and the larger biology community) assess models using data-driven tests. 
\textit{SciUnit} can be applied similarly to other neuroscience (or biology) sub-disciplines using \textit{NeuronUnit} analogues written by the corresponding communities.

@@

\subsubsection{Capabilities from NeuroTools}
NeuroTools (\url{http://neuralensemble.org/NeuroTools}) is a Python library supporting tasks associated with analysis of neural data (or model output), such as membrane potential time series, spike trains, etc. 
It is an open source and actively developed project, containing reliable algorithms on which to base neurophysiology tests.

@@

\subsubsection{Reference Data for Tests from NeuroElectro}
Answering \textit{how well} requires validation testing against data. 
The NeuroElectro project (\url{http://neuroelectro.org}) is an effort to curate all published single cell neurophysiology data \citep{tripathy_neuroelectro:_2012}.  
Currently, up to 27 electrophysiological properties are reported for 93 cell types, spanning $>$ 2000 single pieces of published data extracted from article tables. 
We have made it easy to construct \textit{NeuronUnit} tests using the NeuroElectro API to get reference data. 
Tests can be based upon data from single journal articles, or from ensembles of articles with a common theme (e.g. about a particular neuron type). 
The former is illustrated in Figure \ref{fig:neuronunit_example}. 
Associated statistics of that data (e.g mean, standard error, and sample size) are attached and enable judgement of model output according to a chosen scoring mechanism. 
While NeuroElectro alone cannot judge all model aspects, it can serve to validate basic features of many neurophysiology models, such as resting membrane potential, action potential width, after-hyperpolarization amplitude, etc. 
As NeuroElectro is the only publicly curated source of such data, it represents a key component for \textit{NeuronUnit} test constuction.

\subsection{\textit{SciDash}: A community web application}\label{sec:scidash}
A collection of tests, models, and test records can summarize the current state of modeling in a research area.  
Community-oriented cyberinfrastructure to support the creation and summarization of such collections is essential.
We used a service called \textit{SciDash} (\url{http://www.scidash.org}) for coordinated development of software repositories.

@@

\textit{SciDash} exists in three layers.
The \textbf{first layer} is completely under the control of individual developer communities, and consists of git repositories \citep{ram_git_2013} stored on GitHub \url{http://github.com}. 
\textit{SciDash} identifies all similarly structured repositories (by tracking the lineage of a vanilla \textit{SciDash} repository above using the GitHub API).
This list of repositories is indexed by \textit{SciDash} to form the \textbf{second layer}, an overview of the state of all \textit{SciDash} repositories on GitHub.  
This index is searchable and filterable on the \textit{SciDash} website to identify repositories of interest to a particular research community.  
The \textbf{third layer} is a ``dashboard'' view of the state of a field, visible through a matrix of records for model/test combinations in a GitHub repository (Fig. \ref{fig:scidash_repo}). 
Test records are visualized as a ``record matrix'' composed of large numbers of model/test combinations (Fig. \ref{fig:scidash_matrix}).  
Each row in this matrix contains results for all tests taken by one model and serve as clear evidence of that model's scope and validity.  
Hyperlinks in the record matrix point to models, tests, and records available on GitHub. 
Each record links to test records (stored in the repository), displaying line-by-line the executed code and intermediate output statements, as well as the underlying models and tests.

@@

\subsection{A Complete Pipeline}
Although the tools described in Sec. \ref{sec:neuronunit_acitivities} do not exhaust the possible sources of models, capabilities, and test data, they provide an immediate point of entry into the neurophysiology community and a powerful demonstration of our proposal. 
In the \textit{NeuronUnit} repository (\url{http://github.com/scidash/neuronunit}) is a runnable script (\textit{examples.py}) demonstrating a complete testing pipeline. 
It (1) selects an OSB model; 
(2) simulates it using NeuroConstruct; 
(3) tests the widths of the resulting action potentials, extracted and computed using NeuroTools, against NeuroElectro data downloaded on-the-fly, using a \textit{NeuronUnit} test class called \verbx{SpikeWidthTestDynamic}; 
and (4) computes and prints a test score (Figure~\ref{fig:neuronunit_example}).

@@

\subsection{Participation from Modeling Communities}
Modelers may not want to expose model capabilities, a requirement of test-taking.  
We anticipate four solutions: \textbf{First}, interfacing a model to \textit{SciUnit} requires only implementing selected model capabilities. 
Often this means identifying native model procedures that satisfy a capability, and wrapping their functionality. 
This can require as little as one line of code. 
Importantly, the modeler is not required to expose or rewrite any model flow control. 
\textbf{Second}, we support multiple environments automatically by using NeuroML \citep{gleeson_neuroml:_2010}, and other simulator-independent model descriptions are possible for other domains. 
Automated generation of NeuroML from native model source code is in development (Gleeson, personal communication); for the popular NEURON simulator (\url{http://www.neuron.yale.edu/neuron}), this functionality is already mature and in use. 
This minimizes modeler effort for a large and growing number of models. 
\textbf{Third}, modelers have an incentive to demonstrate publicly their models' validity. 
Participation in public modeling competitions demonstrates this incentive. 
\textbf{Fourth}, modelers have an incentive to use \textit{SciUnit} during development (see TDD, above) to ensure that ongoing development preserves correspondence between model and data. 
A popular test suite can represent a ``gold standard'' by which progress during development is judged.

@@

\subsection{Diversity of Levels and Kinds of Models and Data}
The diversity of topics in biology is vast. 
\textbf{First}, we address this by providing an interface allowing modelers to express specific capabilities.  
This capability set determines the range of eligible tests.  Scale hierarchies are embedded in capability inheritance.  =
For example, \verbx{HasActionPotentials} inherits from \verbx{HasNeurons}, and \verbx{HodgkinHuxley} inherits from \verbx{VoltageGated}. 
Thus, incompatibility of a test-requiring-action-potentials for a model-lacking-neurons is known without explicit tagging. 
\textbf{Second}, NeuroML naturally addresses diversity of scales because it is organized hierarchically, in ``levels.''  
Models can be sub- or supersets of other models; similarly for SBML (\citep{hucka_systems_2003}, \url{http://sbml.org}) a general systems biology markup language. 
\textbf{Third}, cross-level testing can use ``Representional Similarity Analysis'' (RSA) \citep{kriegeskorte_representational_2008}, requiring only that a model respond to defined inputs (e.g. stimuli). 
A ``similarity matrix'' for input responses defines a unique model signature, and can serve as intermediate test output. 
Goodness-of-fit between similarity matrices for model and experiment determines test scores; 
these matrices are independent of model scale because their size depends only on test inputs, not system detail.

\subsection{Arbitrary Scoring Criteria for Tests}
A test first assesses goodness-of-fit, and applies a normalization (e.g. pass/fail, 0.0-1.0) to generate a score. 
Arbitrary choices at both stages may benefit some models over others.  \textbf{First}, however, rank-ordering is constant across many goodness-of-fit metrics, meaning that choice of metric will rarely cause an otherwise passing model to fail and vice versa. 
For example, given a data mean and variance, ordering model output by Z-score or p-value will yield the same relative ranking of models. 
Indeed, rank ordering of models may prove more valuable than test scores themselves. 
\textbf{Second}, suite repositories are open (e.g. Fig. \ref{fig:scidash_repo}), so tests can be cloned and new statistical choices implemented. 
Statistics as applied to neuroscience have been notoriously ``fast and loose''; 
identification and correction of flawed methodology is becoming increasingly common (\citep{button_power_2013,kriegeskorte_circular_2009,galbraith_study_2010}, \url{http://prefrontal.org/files/posters/Bennett-Salmon-2009.jpg}), and is accelerated by an open framework. 
The validation criteria is made explicit in the specification of a test, so modelers need not guess which criteria are being used to validate or invalidate their model. 
Validation criteria are subject to debate (indeed, the QSNMC criteria changed between 2007 and 2008 due to such debates), and neuroscientists who wish to promote different criteria need only derive alternative tests. 
A community consensus will slowly emerge in the form a commonly-accepted test suite (see Sec. \ref{sec:scidash}). 
This community can judge which test version is most appropriate, i.e. what a model \textit{should} do -- this process documented via standard moderation techiniques used on GitHub -- and the \textit{SciUnit} framework determines whether the model \textit{does} it.
